---
title: "Prediction Assignment"
author: "Karin Kolbe"
date: "24/12/2020"
output: html_document
---
## Summary
Subjects were ask to perform barbell lifts correctly and incorrectly in 5 different ways. See http://groupware.les.inf.puc-rio.br/har.
This assignment is to use the Test data to build a prediction model to accurately predict a set of 20 entries.

The approach used is to split the Data into testing and training. Then apply two different models, a Decision Tree and Random Forrest. The latter is much more accurate and is used to apply to the Validation sample.

#### Packages
```{r message=FALSE}
library(dplyr)  
library(knitr)

library(randomForest)  
library(caret)
library(rattle)
library(gbm)
library(rpart)
```

```{r}
setwd("~/dataScience/8 Practical Machine Learning/Assig")
rm(list=ls())
```


## Download and read data
```{r download, cache=TRUE}

#Function to download & read files
getData <- function (SrcFile, SrcURL){
        download.file(url = SrcURL, destfile = SrcFile, method = "curl")
        read.csv(SrcFile ,na.strings=c(" ", "", "#DIV/0!", "NA"))
        }

Training0 <- getData("pml-training.csv",
        "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
Finaltest <- getData("pml-testing.csv",
        "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")

print(dim(Training0))
print(dim(Finaltest))
```

## Tidy the data
```{r}
#Remove columns X,  user_name, timestamps, & window info

Training1 <- Training0[,-c(1:7)]
print(dim(Training1))
#Remove NA rows
Training2<-Training1[,!apply(Training1,2,function(x) any(is.na(x)) )]
print(dim(Training2))
```

## Split data into training and testing
```{r}
set.seed(679)
splitter<-createDataPartition(Training2$classe, p=0.6, list =FALSE)
train <- Training2[splitter,]
test <- Training2[-splitter,]
print(dim(train))
print(dim(test))
```


## Decision Tree
```{r}
set.seed(567)
model_DT<-rpart(classe ~ .,data=train,method="class")

```

```{r}
fancyRpartPlot(model_DT)
```

Now apply to Test data
```{r}
predict_DT <-predict(model_DT,newdata=test,type="class")
conMatrix_DT <- confusionMatrix(predict_DT, test$classe)
conMatrix_DT
```

Accuracy is low at 74.31% so try a different approach.

### Random Forest

Use with cross validation
```{r}
set.seed(789)
model_RF <- randomForest(classe~., data=train, method="class",
                        trControl = traincontrol(method="cv"),number=3)
model_RF
```

```{r}
predict_RF <- predict(model_RF,test, type='class')
conMatrix_RF <- confusionMatrix(predict_RF,test$classe)
conMatrix_RF
```

## Out of sample error rate 
For the Decision Tree the accuracy was 74.3% and Random Forest is 99.4%, thus use this for the Validation 


# Validation
Apply the Random Forest model to the Validation data
```{r}
TestResult <- predict(model_RF,Finaltest, type="class")
TestResult
```


